Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2017-03-07 ms_pre_post.Rmd
Most of the Discussion is now edited, though needs to be reviewed. Last section remains

@@

@@

<!--- I actually don't understand the motivation behind any of these suggested changes, nor the internal logic of some of these sentences. The proposed first sentence of the introduction now shifts from talking generally about all of our results to only referring to the first stated question. The proposed second sentence is especially difficult for me -- I'm not sure how to start editing it. The second sentence suggests that evolution of degenerate traits can't be assumed to occur independently (but this is trivially true -- almost by definition! And it inverts the key logic -- the point is that traits don't have independent effects *potentially leading to degeneracy*). The next part of the second sentence suggests it necessarily follows from degeneracy that relative costs will shape reproductive interactions, but why does this have to be the case? There needs to be some logical connection between these two claims that just isn't there. I can't really figure out how to edit this because I'm really unclear as to what the logical progression of ideas is here. Can you clarify what you mean by this? --->

Evolution of reproductive strategies cannot be presumed to occur in isolation in wild populations, meaning that observed patterns of reproductive strategy within and across taxa are likely to be affected by evolutionary feedbacks and (potentially asymmetric) costs. We have addressed key questions regarding the effect of such feedbacks and costs on evolution of pre-copulatory inbreeding avoidance, post-copulatory inbreeding avoidance, and polyandry. A critical analysis of our model engenders new predictions concerning evolutionary dynamics given the joint evolution of these traits, which may be used to formulate new empirical hypotheses within and across taxa.

For example, our results highlight a fundamental but theoretically under-developed relationship between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance. Post-copulatory inbreeding avoidance is useless for monandrous females, and most effective for highly polyandrous females that can choose amongst multiple mates to ensure that offspring are fertilised favourably. In contrast, pre-copulatory mechanisms are most salient for monandrous females that have no post-copulatory means for avoiding inbreeding. This theory is borne out in our simulation results, as selection for post-copulatory inbreeding avoidance was negligible in monandrous populations, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). The effect of monandry on selection for pre-copulatory inbreeding avoidance was weaker, but pre-copulatory inbreeding avoidance was generally stronger when populations where monandrous (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry increasing selection for post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

@@

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population by modifying its effect on inbreeding avodance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and monandry is common [@Lihoreau2007; @Metzger2010; @Metzger2010a], evolution of post-copulatory inbreeding avoidance is highly unlikely.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g., @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (*Poecilia reticulata*) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids [@Gasparini2011]. In black field crickets (*Teleogryllus commodus*), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs [@Bussiere2006; @Tuni2013]. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for further polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

Conclusion
-------------------------------------------------------------------

Evolution of reproductive strategies involves multiple potentially interacting and degenerate phenotypes. Here we modelled evolution of three reproductive strategies including polyandry, pre-copulatory inbreeding avoidance, and post-copulatory inbreeding avoidance, focusing on three key questions. Using individual-based modelling, we found that evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry, that evolution of a costly inbreeding avoidance strategy was inhibited when the other inbreeding avoidance strategy was cost free, and that evolution of post-copulatory inbreeding avoidance but not pre-copulatory inbreeding avoidance was precluded when the other inbreeding avoidance strategy was fixed for strong inbreeding avoidance. We thereby conclude that evolution of inbreeding avoidance strategies and associated polyandry are likely to interact in predictable ways, potentially generating new hypotheses concerning evolution of reproductive strategies within and across taxa.

>>>>

<!--- ==================================================================== --->
<!--- BEGIN COMMENTS NOT TO BE TRANSFERRED TO THE WORD DOCUMENT FOR COAUTHORS
<!--- ==================================================================== --->

<!--- Evolution of reproductive strategies involves multiple potentially interacting and degenerate phenotypes. Here we modelled evolution of three reproductive strategies including polyandry, pre-copulatory inbreeding avoidance, and post-copulatory inbreeding avoidance, focusing on three key questions. Using individual-based modelling, we found that evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry, that evolution of a costly inbreeding avoidance strategy was inhibited when the other inbreeding avoidance strategy was cost free, and that evolution of post-copulatory inbreeding avoidance but not pre-copulatory inbreeding avoidance was precluded when the other inbreeding avoidance strategy was fixed for strong inbreeding avoidance. We thereby conclude that evolution of inbreeding avoidance strategies and associated polyandry are likely to interact in predictable ways, potentially generating new hypotheses concerning evolution of reproductive strategies within and across taxa. --->

@@

<!--- Too much/too vague no doubt – but this is a hanger to link back to the bit you flagged up as a worry in the Intro, about whether someone would query why we need ‘dynamic models’ – so here we could flag up the dynamicism again somehow? --->

@@

<!--- I was wondering about reiterating a clearer KG here, but actually I think it’s fine without – see comment attached to end of this suggested paragraph. --->

<!--- One of my main comments about the Discussion at the moment is that the really clear structure of the 3 questions that’s now set up at the end of the Intro and followed through the Methods and Results has completely disappeared here. Can we get it back again, a bit? Restating the questions here might be overkill, but yet I’m thinking that a short summary version might be helpful – something like this? --->

<!--- I didn’t get much from this here, and anyway you pretty much repeat the exact same thing in the ‘conclusions’ section at the end? So, if we have to have a conclusions section (for JEB?), then I think it’s best there? --->

<!--- This sentence might need tweaking a bit, but I think something like it would be a good strong end to this first para? --->

<!--- So this is where I start to find the structure of the Discussion a bit hard to follow – seem to be addressing the three questions in reverse order. This first bit (and the link to fig 4), is really addressing question 3, isn’t it? And then we go on to Q2 (fig 3) and then finally Q1 (fig 1). This is pretty confusing. Would it be possible to turn this section around so that the three questions are dealt with in the ‘right’ order? I haven’t tried to implement. Or if you want to keep it like this we might need to clarify that structure and (implicitly or explicitly) note why things are reversed? --->

<!--- ==================================================================== --->
<!--- END COMMENTS NOT TO BE TRANSFERRED TO THE WORD DOCUMENT FOR COAUTHORS
<!--- ==================================================================== --->

Evolution of reproductive strategies cannot be presumed to occur in isolation in wild populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently feed back to drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be affected by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

<!--- I have tried to re-work this first paragraph, but I had a very hard time following the logical flow of ideas given the suggested changes. I understand that you want to get all of the components of this manuscript into the front of the Discussion, but I think they need to fit together much more coherently. I didn't quite understand the motivation behind the suggested changes. For example, the proposed first sentence of the Discussion shifted from talking generally about all of our results to only referring to one of our questions. I didn't understand the proposed second sentence at all, which suggested that evolution of degenerate traits can't be assumed to occur independently? (But this is trivially true -- almost by definition! And it inverts the key logic -- the point is that traits don't have independent effects *potentially leading to degeneracy*).  The next part of the proposed second sentence suggested that it necessarily follows from degeneracy that relative costs will shape reproductive interactions, but why does this have to be the case? There needs to be some logical connection between these two claims that just isn't there. I couldn't really figure out how to re-work it because I was really unclear as to what the logical progression of ideas was. Can you clarify what you mean by this? --->

<!--- I think the logic of the first sentence is important here, though it might need some tweaks. The point is that phenotypes don't occur in isolation -> multiple phenotypes interact -> two types of interaction are feedbacks and degeneracy. Hence, degeneracy is one *type* of way that phenotypes might interact, so I think that the sentence structure needs to reflect this so readers don't get confused (i.e., we can't just skip to 'such degenerate' because it doesn't clearly follow from the previous sentence. --->

<!--- I don't think just mentioning 'dynamic models' is enough to make a convincing argument that they were necessary to address the knowledge gap. If we really wanted to flesh out the idea, we would have to explain *why* we couldn't get the answer we did with non-dynamic models. This isn't obvious or trivial! --->

<!--- I have re-ordered the Discussion to reflect the order of the questions, but this sacrifices something by putting the specifics points ahead of more general questions (i.e., those references to Figure 4 were not related to any of the three questions specifically, but were used to help make a more general point derived from a holistic understanding of the model). Hence, I have created a new section for more general points concerning reproductive strategy, also removing the conclusion. --->

We predict evolution of polyandry in populations where inbreeding depression is severe and inbreeding avoidance through post-copulatory mechanisms can also evolve. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g., @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (*Poecilia reticulata*) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids [@Gasparini2011]. In black field crickets (*Teleogryllus commodus*), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs [@Bussiere2006; @Tuni2013]. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for further polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population by modifying its effect on inbreeding avodance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs thereby reflect empirical observations in some populations and are therefore biologically realistic, other types of costs have also been observed and could therefore be considered in future models. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

<!--- I don't understand the contrast between the two comments, the first of which suggests that the empirical examples of costs cited are not needed here, and the second of which suggests that we need to be more explicit about why we did things the way we did. Hence, I have revised to clarify that we made the modelling choices that we did with respect to costs *because they are observed in empirical systems*. Of course, the costs that we modelled are not the *only* kinds of 'costs' that are observed in empirical systems (hence the Roberts reference), so other models might want to look into different types of costs. I agree that it would not be at all surprising if a reviewer raised a question about how we modelled the costs (or anything else), but this always seems to be true for modelling and I think it kind of misses the big picture about how models like this try to develop theory. We can't model *every* kind of cost that we can possibly think of, and we're not modelling a specific system where we might be able to justify some costs being more reasonable than others. We also can't (I don't think) logically show with an IBM that the type of cost will be arbitray with respect to the questions of interest (i.e., any cost will do). In the end, we have to make reasonable decisions about how to model costs and work from there. I think that citing empirical studies shows that our assumptions of cost type are reasonable for some systems (though potentially not others). Had we made different assumptions, I suspect we'd get queries from some readers wondering why we didn't do it a different way. --->

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and polyandry is uncommon [@Lihoreau2007; @Metzger2010; @Metzger2010a], evolution of post-copulatory inbreeding avoidance is highly unlikely.

General predictions concering inbreeding avoidance and polyandry
-------------------------------------------------------------------

Post-copulatory inbreeding avoidance is useless for monandrous females, and most effective for highly polyandrous females that can choose amongst multiple mates to ensure that offspring are fertilised favourably. In contrast, pre-copulatory mechanisms are most salient for monandrous females that have no post-copulatory means for avoiding inbreeding. This theory is borne out in our simulation results, as selection for post-copulatory inbreeding avoidance was negligible in monandrous populations, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). The effect of monandry on selection for pre-copulatory inbreeding avoidance was weaker, but pre-copulatory inbreeding avoidance was generally stronger when populations where monandrous (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry increasing selection for post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

<!---  Diluting the idea of degeneracy throughout the manuscript now makes me worry that some readers will miss it entirely. If they're not reading carefully, they'll miss (or simply forget) the buried definition in the fourth paragraph of the Introduction. Because the concept of degeneracy becomes so important later on, they'll need to back track to remember what it actually means. I'm less a fan of 'flagging' things for this reason -- readers don't really *need* to understand degeneracy completely until now, but here they have to go all the way back to the Introduction to get the full scope of its importance. --->

@@
